Speculation in foreign exchange: noise or information

### 2015-06-05 Knitr/NewPaper.Rnw
Updated the first part of the methods. This needs to be looked at and corrected.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.  
    The foreign exchange market is utilised because it is the largest and most liquid market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}. In addition, the cost of trading in this markket, when measurd in terms of the size of the bid-ask spread, is extremely low. An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}.

@@

\bibliography{../../../../myrefs}

>>>>

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.  
    The foreign exchange market is utilised because it is the largest and most liquid market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}. In addition, the cost of trading in this markket, when measurd in terms of the size of the bid-ask spread, is extremely low. An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}. The idea of noise trading may be particularly important for the foreign exchange market where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citep{FrootChartists} for a survey.

@@

\section{Methodology}

The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculators' activity in the  foreign exchange market .   This will involve three steps:  first,assess the price implication of the noise-trader modele when speculative sentiment or activity becomes extreme; second, measure speculative activity in the foreign exchange market; finally,  whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.

The noise-trader model used in this study is from De Long, Shleifer, Summers and Waldeman \citep{Delong1990noise}. This is a derivative of the Samuelson overlapping generations model  with two-period agents making a portfolio choice \citep{Samuelson1958}.  There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.  There are two types of investor or trader: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

\begin{equation}
\rho_t \sim N(\rho^*, \sigma_{\rho^2})
\end{equation}

where $\rho^*$ is the average mistake of the noise-traders and $\sigma_{\rho}^2$ is the variance of the misperception.      Assuming an exponential utility function or constant absolute risk aversion\footnote{Though \citep{TobinLiquidity} used a quadratic utility function to show that investor preference could be defined by just the mean and the variance of the returns, the increasing absolute risk aversion this implies does not seem to be a suitable property as it implies people become more unwilling to take risk as they become more wealthy (\citep{Hicks1962} and \citep{ArrowRisk}.   Constant absolute risk aversion is the \citep{PrattCARA}, \citep{ArrowRisk}  measure based on $u(c) = 1 - e^{-\gamma c}$ where the \emph{coefficient of absolute risk aversion} $A(c) = -\frac{u''(c)}{u'(c)} = \gamma$.  This is a standard device.}, each type of investor chooses the proportion $(\lambda)$ of the risky asset to hold to maximise utility.

\begin{equation}
E(U) = \bar{w} - \gamma \sigma_{w}^2
\end{equation}

where $\bar{w}$ is end of period wealth, $\gamma$ is a measure of absolute risk aversion and $\sigma_w^2$ is the variance of wealth. Utility is a function of risk and return.  This depends on holdings of the risky asset.   For the informed trader this is,

\begin{equation}
E(U) = c_0 + \lambda_t^i[r + _tp_{t+1} - p_t(1 + r)] - \gamma (\lambda_t^i)^2 \left (_t\sigma^2_{p, t+1} \right ) 
\end{equation}

where $c_0$ is the endowment, following the notation in the original paper, $_tp_{t+1}$ is the expected price of the risky asset at $t + 1$, conditional on information at time $t$ and $_t\sigma^2_{\rho, t+1}$ is the expected variance of the misperception conditional on information at time $t$.

Maximising utility with respect to holdings of the risky asset and iterating backward to solve for the current value,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.

\begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma}{r}\ \big (_t\sigma^2_{p,t+1} \big )
\label{eq:NP1}
\end{equation}

In words, the deviation from the true fundamental value of unity is a result of the discounted value of the noise-traders' misperception of the price $(p_t - \rho^*)$, multiplied by the weight of noise-traders in the total $(\mu)$ plus a perpetuity of the average noise-trader misperception $(\frac{ \mu \rho^*}{r} )$, adjusted for the risk.   This risk is a function of the volatility of the price which,  from equation \ref{eq:NP1} is a result of the variability of the misperception of speculative noise-traders.   All but the second term in equation \ref{eq:NP1} are constants, and if the variance of the price one period ahead  is equal to the variance of the misperception,

\begin{equation} 
 _t\sigma^2_{p,t+1} = \frac{\mu^2 \sigma_{\rho}^2}{1+r^2}  
\end{equation}

the final price equation becomes.

\begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r)^2}
\label{eq:NP2}
\end{equation}

The paper makes use of the utility function to compare the outcomes for the two types of investor for different realisations of the parameters.  They note four effects:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief.  The Friedman Effect and Price Pressure tend to reduce noise-trader utility as they work to create losses or reduced gains for noise-trader positions; Hold More and Create Space can increase noise-trader utility and profits \citep[pp. 14-15]{Delong1990noise}. 
%Put something on these four in the conclusion. 
A number of important ideas flow from the model.  The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citep{FriedmanPositive}.  While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.  The noise-traders expected to get higher utility due to their misperception.  The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.  This is consistent with the type of over-confident, entrepreneurial misperception discussed in Section \ref{sec:prob} in relation to the returns to speculation.  If the noise-traders become aware of the risk that they are taking they are less likely to undertake the activity.

De Long and the others test a variety of specifications that include modest switching from one style of investment to another and changes in the weight of new traders after previous performance.  Some result in the disappearance of noise-traders while others suggest that they can become more dominant.  The model also explains the importance of investment with a long-term horizon as this is the only way that \emph{noise- trader risk} can be overcome.  If agents live for more than two periods, fundamental investors can make an arbitrage between safe and risky assets, driving the price back to fundamental value.  Therefore, investors or traders with a reputation that will allow resources and time to consider the long-term would be able to take full advantage of the noise-traders' misperception.  The model relies crucially on the institutional or psychological myopia. 
% read again and add more.

Noise-traders drive the price in the direction of their mis-perception.    The second and third terms in Equation \ref{eq:NP2} show the deviation from fundamental value due to the fact that the noise-trader mis-perception is not equal to zero but equal to $p^*$.  This means that the mis-perception is not random but biased in a positive fashion.  If the mis-perception is random, we are back to the random walk with noise-traders adding liquidity on each side of the market and the informed traders able to take advantage of any deviation of the price from its fundamental value.  There is still some risk that the outcome will be different from the fundamental value, because the noise-traders may still become more extreme in their misperception, but on average informed traders will make profit if they act.  This deviation in the price is what De Long and co-authors call price pressure.  While there is no fundamental risk in this model, price pressure causes risk for which there is no compensation and which is assumed by the noise-traders.  This is rather similar to the risk created by speculators in the Keynes models where additional uncertainty, in the Knightian sense, encourages more investors towards self-perpetuating myopic behaviour.

%One of the key assumptions of the model is the two period construction that means that final value is not determined.  This means that informed investors cannot just wait for the fundamental value to assert itself.  Therefore, the model is consistent with the sort of myopia that may be caused by impatience \citep{Keynes1936}, funding costs, social pressure and institutional features. % references needed.  
%\footnote{An good practical example is that of Tony Dye the Philips and Drew Fund Manager who refused to participate in the technology boom of the late 1990s.  The relative under-performance of the fund cause a loss of investors and his eventual resignation in March 2000.  This is the same date as the peak of the US technology index NASDAQ.  It is also consistent with the words of Citicorp Chief Executive Charles Prince who is reported to have said, "As long as the music's playing, you've got to get up and dance" when asked by the \emph{Financial Times} in July 2007 about the bank's lending activity.}

%is this paragraph necessary?
%The De Long and others do not go into the details of the misperception that is suffered by the noise traders.  There are a number of explanation for this deviation of the misperception from zero.  The first are the behavioural biases that were addressed in section \ref{sec:behavioural}.   If investors generally are suffering from the \emph{representativeness} heuristic an over-focus on the immediate information that the expense of the bigger picture or the base risk to securities value would cause a systematic deviation in favour or optimism or pessimism.  This effect can be exacerbated by the overconfidence that has been identified generally in the behavioural economic literature and by Odean specifically for investment markets.  Systematic bias creates price pressure which feeds back into a successful outcome that is regarded as being the consequence of skill, ability and good judgment and investment talent rather than luck and temporary circumstance.  %more may be required here Hirshliefer and Luo (2001) on the survival of noise traders.
%There is also the social pressure and a feedback mechanism that comes from taking an action that others are taking.  Therefore, if others have the same investment strategy, this can make investors more confident and, in some cases, will cause investors to infer private information from the actions of other investors who are basing investment decisions on the same style of imitation.  All the investors think that they are in good company and that the knowledge of the others fills the gaps where their own uncertainty begins. % need references for this.

\
\bibliography{../../../../myrefs}
